Too hard to get: the role of probabilistic expectations and cognitive complexity in destructive multi-dimensional reference points

### 2025-02-27 manuscript-apa.qmd
Aspen Han - Project submission #1 for grading - @nrdowling @lim1an

format:
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    fig-width: 6
    fig-asp: 0.618
    fig-align: center
    include-in-header:
      text: |
        \usepackage{enumitem}
        \setlist[itemize]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \setlist[enumerate]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
---

@@

$$U = p(e_1, e_2) - c(e_1, e_2) +
  μ_1[(n(e_1 )-N)\mathbb{I}(n<N) + λ_1(n(e_1)-N)\mathbb{I}(n≤N)] +
  μ_2[(Q-q(e_2))\mathbb{I}(q<Q) + λ_2(Q-q(e_2))\mathbb{I}(q≥Q)]$$

@@

$$U = p(e_1, e_2) - c(e_1, e_2) + \\
  E[\mathbb{P}(\{n≥N-ε\}∩\{q≤Q+ε\}] \times \theta \times \\
  \{μ_1[(n(e_1 )-N)\mathbb{I}(n<N) + λ_1(n(e_1)-N)\mathbb{I}(n≤N)] + \\
  μ_2[(Q-q(e_2))\mathbb{I}(q<Q) + λ_2(Q-q(e_2))\mathbb{I}(q≥Q)]\}$$

The first additional term $E[\mathbb{P}(\{n≥N-ε\}∩\{q≤Q+ε\}]$ captures the agent’s expected probability of simultaneously achieving (within some bandwidth $\varepsilon$ of) all reference points. When this expected probability is lower, the agent weights the gain-loss utilities less and hence is less responsive to the reference points. The second additional parameter $\theta≥0$ is a parameter decreasing in the cognitive complexity required to integrate the multiple reference points, so greater cognitive complexity attenuates reference point effects.

@@

KR model prediction:

KR1. Treatments 2 and 4 will have similar positive effects on the probability of achieving any target.
KR2. Treatment 3 should have a larger positive effect than treatment 2 (and 4) for achieving both targets or for achieving $Q$ but not $N$.

@@

A1. Treatment 3 should have lower positive effect for achieving any target than treatment 2.
A2. Treatment 4 should have a lower positive effect for achieving any targets than treatment 2.

>>>>

format:
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    fig-width: 6
    fig-asp: 0.618
    fig-align: center
    include-in-header:
      text: |
        \usepackage{amsmath}
        \usepackage{enumitem}
        \setlist[itemize]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
        \setlist[enumerate]{topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt}
---

@@

\begin{align}
U = &p(e_1, e_2) - c(e_1, e_2) + \nonumber \\ 
    &\mu_1[(n(e_1 )-N)\mathbb{I}(n>N) + \lambda_1(n(e_1)-N)\mathbb{I}(n \leq N)] + \nonumber \\
    &\mu_2[(Q-q(e_2))\mathbb{I}(q<Q) + \lambda_2(Q-q(e_2))\mathbb{I}(q \geq Q)] \nonumber
\end{align}

@@

\begin{align}
U = &p(e_1, e_2) - c(e_1, e_2) + \nonumber \\
    &E[\mathbb{P}(\{n \geq N-\varepsilon \} \cap \{q \leq Q+ε\}] \times \theta \times \nonumber \\
    &\{\mu_1[(n(e_1 )-N)\mathbb{I}(n>N) + \lambda_1(n(e_1)-N)\mathbb{I}(n \leq N)] + \nonumber \\
    &\mu_2[(Q-q(e_2))\mathbb{I}(q<Q) + \lambda_2(Q-q(e_2))\mathbb{I}(q \geq Q)]\} \nonumber
\end{align}

The first additional term $E[\mathbb{P}(\{n \geq N-\varepsilon \} \cap \{q \leq Q+ε\}]$ captures the agent’s expected probability of simultaneously achieving (within some bandwidth $\varepsilon$ of) all reference points. When this expected probability is lower, the agent weights the gain-loss utilities less and hence is less responsive to the reference points. The second additional parameter $\theta≥0$ is a parameter decreasing in the cognitive complexity required to integrate the multiple reference points, so greater cognitive complexity attenuates reference point effects.

@@

KR model predictions:

* Treatments 2 and 4 will have similar positive effects on the probability of achieving any target. --- (KR1)
* Treatment 3 should have a larger positive effect than treatment 2 (and 4) for achieving both targets or for achieving $Q$ but not $N$. --- (KR2)

@@

* Treatment 3 should have lower positive effect for achieving any target than treatment 2. --- (A1)
* Treatment 4 should have a lower positive effect for achieving any targets than treatment 2. --- (A2)
